A quantum parallel parallel Markov chain Monte Carlo

### 2021-11-12 main.tex
gumbel max correct
grover algortithm

% !TeX document-id = {d42653b9-9ea4-48c2-860b-3d6e6e9d8ab0}
% !TeX TXS-program:compile = txs:///pdflatex/[--shell-escape]
\documentclass[12pt]{article} % For LaTeX2e
%\usepackage{neurips_2021}
\usepackage[colorlinks, citecolor={blue}]{hyperref}
\usepackage{url}
\usepackage{amsfonts,amscd,amssymb}
\usepackage{amsthm,amsmath,natbib}
\usepackage{algorithm,algorithmicx,algpseudocode}
\usepackage{bm}
\usepackage{bbm} %bb font numbers
\usepackage[table]{xcolor}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{natbib}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
%\usepackage[nolists]{endfloat}
\usepackage{listings}
\usepackage[textsize=tiny]{todonotes}
\usepackage{tikz}
\usetikzlibrary{shapes.misc}
\usepackage{etoolbox}
\usepackage{appendix}
\usepackage[format=plain,
labelfont={it},
textfont=it]{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{xr}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{authblk}
\usepackage{mathbbol}
\usepackage{braket}

@@

@@

@@

@@

We wish to randomly select a single element from the set $\{0,1,\dots,P\}$ with probability proportional to the unnormalized probabilities $\ppi^*=(\pi_0^*,\pi_1^*,\dots,\pi_P^*)$.  That is, there exists a $c>0$ such that $\ppi^* = c \ppi$, for $\ppi$ a valid probability vector, but we only have access to $\ppi^*$.  Define $\llambda:= \log \ppi$ and $\llambda^*:= \log \ppi^*=\log\ppi + \log c$,  and assume that $z_0,\dots,z_P\stackrel{iid}{\sim}Gumbel(0,1)$. Then, the probability density function $g(\cdot)$ and cumulative distribution function $G(\cdot)$ for each individual $z_p$ are
\begin{align}
g(z_p) = \exp\big(-z_p+ \exp(-z_p)\big) \, , \quad \mbox{and} \quad G(z_p) = \exp \big( -\exp(-z_p) \big)\, .
\end{align}
Now, defining the random variables $\alpha^*_p:=\lambda^*_p+z_p$, $\alpha_p:=\lambda_p+z_p$ and  
\begin{align}
\hat{p} = \argmax_{p=0,\dots,P}\: \alpha^*_p \, ,
\end{align}
we have the result
\begin{align}
\mbox{Pr} (\hat{p} = p) = \pi_p \, .
\end{align}
In words, the procedure of adding Gumbel noise to unnormalized log-probabilities and taking the index of the maximum produces a random variable that follows the discrete distribution over $\ppi$.  Moving from left to right:
\begin{align*}
\mbox{Pr} (\hat{p} = p) &= \mbox{Pr} (\alpha^*_p > \alpha^*_{p'}, \, \forall p' \neq p ) = \mbox{Pr} (\alpha_p > \alpha_{p'}, \, \forall p' \neq p ) \\
&= \int_{-\infty}^\infty \prod_{p'\neq p} \mbox{Pr} (\alpha_p > \alpha_{p'}| \alpha_p) \p(\alpha_p) \, \dd \alpha_p 
= \int_{-\infty}^\infty \prod_{p'\neq p} \mbox{Pr} (z_{p'} \leq \alpha_p-\lambda_{p'}| \alpha_p) \p(\alpha_p) \, \dd \alpha_p \\
&=  \int_{-\infty}^\infty  \prod_{p'\neq p} \exp\big( -\exp(\lambda_{p'}-\alpha_p)\big) \exp\big(-\alpha_p+\lambda_p + \exp(-\alpha_p+\lambda_p) \big) \, \dd \alpha_p \, .
\end{align*}
Recalling that $\lambda_p=\log \pi_p$, we exponentiate the logarithms, and the integral becomes
\begin{align*}
& \pi_p \int_{-\infty}^\infty  \prod_{p'\neq p} \exp\big( -\pi_{p'}\exp(-\alpha_p)\big) \exp  \big(-\alpha_p + \pi_p\exp(-\alpha_p) \big) \, \dd \alpha_p \\
&= \pi_p \int_{-\infty}^\infty \exp  (-\alpha_p ) \exp\big( -\sum_{p=0}^P\pi_{p}\exp(-\alpha_p)\big)  \, \dd \alpha_p  \\
&= \pi_p \int_{-\infty}^\infty \exp  (-\alpha_p ) \exp\big( -\exp(-\alpha_p)\big)  \, \dd \alpha_p  = \pi_p \, ,
\end{align*}
where the final equality follows easily from the change of variables $u=\exp(-\alpha_p)$.

>>>>

% !TeX document-id = {d42653b9-9ea4-48c2-860b-3d6e6e9d8ab0}
% !TeX TXS-program:compile = txs:///pdflatex/[--shell-escape]
\documentclass[12pt]{article} % For LaTeX2e
%\usepackage{neurips_2021}
\usepackage[colorlinks, citecolor={blue}]{hyperref}
\usepackage{url}
\usepackage{amsfonts,amscd,amssymb}
\usepackage{amsthm,amsmath,natbib}
\usepackage{algorithm2e}
\usepackage{bm}
\usepackage{bbm} %bb font numbers
\usepackage[table]{xcolor}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{natbib}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
%\usepackage[nolists]{endfloat}
\usepackage{listings}
\usepackage[textsize=tiny]{todonotes}
\usepackage{tikz}
\usetikzlibrary{shapes.misc}
\usepackage{etoolbox}
\usepackage{appendix}
\usepackage[format=plain,
labelfont={it},
textfont=it]{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{xr}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{authblk}
\usepackage{mathbbol}
\usepackage{braket}

@@

\SetKwComment{Comment}{/* }{ */}
\RestyleAlgo{ruled}

@@

\begin{algorithm}[t!]
	\caption{Parallel MCMC}\label{alg:pMCMC}
	\KwData{Initial Markov chain state $\ttheta^{(0)}$; total length of Markov chain $S$; total number of proposals per iteration $P$; routine for evaluating target density $\pi(\cdot)$; routines for drawing random samples from the proposal distribution $Q(\ttheta^{(s)},\cdot)$ and from a $P+1$ discrete distribution $Discrete(\ppi)$ given some probability vector $\ppi=(\pi_0,\dots,\pi_P)$.}
	\KwResult{A Markov chain $\ttheta^{(1)}, \dots, \ttheta^{(S)}$.}
	\For{$s \in\{1,\dots,S\}$}{
		$\sigma_\pi \gets 0$\;
		\For{$p' \in \{0,\dots,P\}$}{
			\uIf{$p'=0$}{$\ttheta_{p'} \gets \ttheta^{(s-1)} $}
			\Else{	$\ttheta_{p'} \gets Q(\ttheta^{(s-1)},\cdot)$\;}
			\For{$p'' \in \{0,\dots,p'-1,p'+1,\dots,P\}$}{
				$q_{p'p''} \gets q(\ttheta_{p'},\ttheta_{p''})$\;
			}
			$\pi_{p'} \gets \pi(\ttheta_{p'}) \prod_{p''\neq p'} q_{p'p''}$\;
			$\sigma_\pi \gets \sigma_\pi + \pi_{p'}$\;
		}
		$\ppi \gets \ppi / \sigma_\pi$\;
		$p \gets Discrete(\ppi)$\;
		$\ttheta^{(s)} \gets \ttheta_p$\;
	}
	\Return{$\ttheta^{(1)}, \dots, \ttheta^{(S)}$}\ .
\end{algorithm}

@@

\begin{algorithm}[t!]
	\caption{Grover's algorithm for quantum search}\label{alg:grover}
	\KwData{An oracle gate $\U_f$ taking $\ket{x}\ket{y}$ to $\ket{x}\ket{y\oplus f(x)}$ for a function $f(x):\{0,\dots,N-1\}\rightarrow \{0,1\}$ that satisfies $f(x_0)=1$ for a single $x_0$; $n+1=\log_2(N)+1$ quantum states initialized to $\ket{0}^{\otimes n} \ket{1}$; an integer $R>\pi\sqrt{N}/4$ denoting the number of Grover iterations.}
	\KwResult{An $n$-bit binary string $x_0$ satisfying $f(x_0)=1$ with error less than $1/N$.}
	$\H^{\otimes n+1}\ket{0}^{\otimes n}\ket{1} \longrightarrow \ket{h}\ket{-}$; \hspace{11.5em} \Comment{\hspace{0.3em}$\ket{h}=\frac{1}{\sqrt{N}} \sum_{x=0}^{N-1} \ket{x}$.} 
	$\ket{h}\ket{-} \longrightarrow \G^R \ket{h}\ket{-} \approx \ket{x_0}\ket{-}$;    \hspace{3em}         \Comment{$\G= \Big(2\ket{h}\bra{h}-\I \Big) \Big( \I - 2 \ket{x_0}\bra{x_0} \Big)$}
	$\ket{x_0} \longrightarrow x_0$\;
	\Return{$x_0$}\ .
\end{algorithm}

@@

We wish to randomly select a single element from the set $\{0,1,\dots,P\}$ with probability proportional to the unnormalized probabilities $\ppi^*=(\pi_0^*,\pi_1^*,\dots,\pi_P^*)$.  That is, there exists a $c>0$ such that $\ppi^* = c \ppi$, for $\ppi$ a valid probability vector, but we only have access to $\ppi^*$.  Define $\llambda:= \log \ppi$ and $\llambda^*:= \log \ppi^*=\log\ppi + \log c$,  and assume that $z_0,\dots,z_P\stackrel{iid}{\sim}Gumbel(0,1)$. Then, the probability density function $g(\cdot)$ and cumulative distribution function $G(\cdot)$ for each individual $z_p$ are
\begin{align}
g(z_p) = \exp\big(-z_p- \exp(-z_p)\big) \, , \quad \mbox{and} \quad G(z_p) = \exp \big( -\exp(-z_p) \big)\, .
\end{align}
Now, defining the random variables $\alpha^*_p:=\lambda^*_p+z_p$, $\alpha_p:=\lambda_p+z_p$ and  
\begin{align}
\hat{p} = \argmax_{p=0,\dots,P}\: \alpha^*_p \, ,
\end{align}
we have the result
\begin{align}
\mbox{Pr} (\hat{p} = p) = \pi_p \, .
\end{align}
In words, the procedure of adding Gumbel noise to unnormalized log-probabilities and taking the index of the maximum produces a random variable that follows the discrete distribution over $\ppi$.  Moving from left to right:
\begin{align*}
\mbox{Pr} (\hat{p} = p) &= \mbox{Pr} (\alpha^*_p > \alpha^*_{p'}, \, \forall p' \neq p ) \\
&= \mbox{Pr} (\alpha_p +\log c > \alpha_{p'} + \log c, \, \forall p' \neq p ) 
= \mbox{Pr} (\alpha_p  > \alpha_{p'} , \, \forall p' \neq p ) \\
&= \int_{-\infty}^\infty \prod_{p'\neq p} \mbox{Pr} (\alpha_p > \alpha_{p'}| \alpha_p) g(\alpha_p-\lambda_p) \, \dd \alpha_p 
= \int_{-\infty}^\infty \prod_{p'\neq p} G(\alpha_p-\lambda_{p'}) g(\alpha_p-\lambda_p) \, \dd \alpha_p \\
&=  \int_{-\infty}^\infty  \prod_{p'\neq p} \exp\big( -\exp(\lambda_{p'}-\alpha_p)\big) \exp\big(-\alpha_p+\lambda_p - \exp(-\alpha_p+\lambda_p) \big) \, \dd \alpha_p \, .
\end{align*}
Recalling that $\lambda_{p'}=\log \pi_{p'}$, we exponentiate the logarithms, and the integral becomes
\begin{align*}
& \pi_p \int_{-\infty}^\infty  \prod_{p'\neq p} \exp\big( -\pi_{p'}\exp(-\alpha_p)\big) \exp  \big(-\alpha_p - \pi_p\exp(-\alpha_p) \big) \, \dd \alpha_p \\
&= \pi_p \int_{-\infty}^\infty \exp  (-\alpha_p ) \exp\big( -\sum_{p'=0}^P\pi_{p'}\exp(-\alpha_p)\big)  \, \dd \alpha_p  \\
&= \pi_p \int_{-\infty}^\infty \exp  (-\alpha_p ) \exp\big( -\exp(-\alpha_p)\big)  \, \dd \alpha_p  = \pi_p \, ,
\end{align*}
where the final equality follows easily from the change of variables $u=\exp(-\alpha_p)$.

\begin{algorithm}
	\caption{The Gumbel-max trick}\label{alg:gm}
	\KwData{A vector of unnormalized log-probabilities $\llambda^*=\log \ppi +\log c$, for $\ppi$ a discrete probability vector with $P+1$ elements.}
	\KwResult{A single sample $p \sim$ \emph{Discrete}$(\ppi)$ satisfying $p\in\{0,1,\dots,P\}$.}
	\For{$p' \in\{0,1,\dots,P\}$}{
		$z_{p'} \gets Gumbel(0,1)$\;
		$\alpha^*_{p'} \gets \lambda_{p'}^* + z_{p'}$\;
	}
	$p \gets \argmax_{p'=0,\dots,P}\alpha^*_{p'}$\;
	\Return{p}\ .

%	$y \gets 1$\;
%	$X \gets x$\;
%	$N \gets n$\;
%	\While{$N \neq 0$}{
%		\eIf{$N$ is even}{
%			$X \gets X \times X$\;
%			$N \gets \frac{N}{2}$ %\Comment*[r]{This is a comment}
%		}{\If{$N$ is odd}{
%				$y \gets y \times X$\;
%				$N \gets N - 1$\;
%			}
%		}
%	}
\end{algorithm}
